---
title: "Structure based drug design"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Download and process starting structure.

Here, we download and clean up the HIV-Pr sturcture. PDB Code: 1HSG from the main PDB database. We will make a show "protein only" and "ligand only" PDB files.

```{r}
library(bio3d)

file.name <- get.pdb("1HSG")
```

We will use `read.pdb()`, `atom.select()`, and `write.pdb()` functions to make our separate "protein-only" and "ligand-only" PDB files.

You "got" the PDB file, now you have to read it to see its contents.
```{r}

hiv <- read.pdb(file.name)

hiv

```

# Prepare the initial protein and ligand input files for AutoDock.

Next, to select the atom look up `?atom.select` then use the correct inputs.
```{r}

prot <- atom.select(hiv, "protein", value = TRUE)
lig <- atom.select(hiv, "ligand", value = TRUE)

prot
lig

```

Now that you've put together the file, write it so you can save "1HSG_prot" and "1HSG_lig" as PDB files.
```{r}

write.pdb(prot, file="1hsg_protein.pdb")
write.pdb(lig, file="1hsg_ligand.pdb")

```

Now, switch to the AutoDock program to open those PDB files.
Change the formats to PDBQT files.
The PDBQT file gives you the atom charge + atom type. 
Now, make a config.txt file with the relevant arguments to specify your files and inputs for the program.

>> Note that we skipped the "prepare the ligand" step during class due to time constraints.

# Prepare a docking configuation file.

Install AutoDock Tools Vina.
Run Vina using the Terminal.
You have to identify WHERE your program is and argue with it. (That's why we made a config. .txt file.)

# Dock the ligand onto the HIV-1 protease.
INPUT > "\Program Files (x86)\The Scripps Research Institute\Vina\vina.exe" --config config.txt --log log.txt

Once the run is complete, you should have two new files all.pdbqt, which contains all the docked modes, and log.txt, which contains a table of calculated affinities based on AutoDock Vina's scoring function. The best docked mode, according to AutoDock Vina, is the first entry in all.pdbqt.

# Read docking results.

Now let's inspect what we just made,
```{r}
result <- read.pdb("all.pdbqt", multi=TRUE)

write.pdb(result, "results.pdb")
```

We can now view the docking as a movie on PyMOL.

# Quantify the RMSD value of the results.

We can use R to quantify the results by calculating RMSD (root mean square distance) between each docking result and comparing it to bio3d crystal structures.

```{r}

num_res <- read.pdb("all.pdbqt", multi = TRUE)
num_res
```

```{r}
ori <- read.pdb("ligand.pdbqt")
ori
```

```{r}

rmsd(ori, num_res)

```

