---
title: "Lecture 6 HW"
author: "Tiffany Luong"
date: "1/31/2020"
output: html_document
---

## Protein Drug Interactions Homework Using Bio3d

```{r}
# Can you improve this analysis code?
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

Q1.
```{r eval=FALSE, include=FALSE}
?read.pdb()
```
read.pdb reads a Protein Data Bank PDB coordinate file.

Q2.
```{r eval=FALSE, include=FALSE}
?trim.pdb
```
trim.pdb produces a smaller PDB object (a subset of smaller atoms) from a large PDB object

Q3.
```{r eval=FALSE, include=FALSE}
?plotb3

plotb3(s2.b, top = FALSE, bot = FALSE)
```

The gray and black rectangles represent the alpha and beta helices. Change top and bot to FALSE to remove the rectangles.

Q4.
A better plot to compare the three proteins would be to plot all three of them on the same graph.
```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```

Q5.
According to the cluster dendrogram, s1.b and s3.b are proteins more similar to each other based on their B-factor trends.

## Q6. Write a function that analyzes protein drug interactions by reading in any protein PDB data and outputs a plot for the specified protein.

Submit as R markdown file and either an HTML or PDF with the code and output saved.

# What are the inputs of the function?
4 main sets of inputs:
1. read.pdb (read the protein pdb file)
2. trim.pdb (reduce the data to a smaller pdb file)
3. chain.pdb (find possible chain breaks based on the Calpha or peptide bond atom separation)
4.plot3b (make a graph of the protein drug interactions from PDB)

Since these are the lines of code that get repeated, it's best to divvy them up into a function.

#What does the function do? How do you use it?
The function **pdi_plot** should take a vector of the PDB files and the parameters to analyze each file (a chain, an element, and a factor). **pdi_plot** will go through the vector and apply the parameters, then creates a single plot for easy visualization of the PDB data.

To use the function, you can fill in the parameters for the vectors PDB files, chains, elements, and factors.

# What is the output of the function?
The output of the function should be a single graph with the three protein-drug-interactions plotted. The color is set to red, blue, and orange but can be customized.

## HOMEWORK CODE

```{r}

library(bio3d)

#change variables to desired
files <- "4AKE"
chains <- "A"
elem <- "CA"

pdi_plot <- function(files) {
  
#input 1: read    
    s <- read.pdb(files)

#input 2: trim
    s.chain <- trim.pdb(s, chain = chains, elety = elem)

#input 3: chain.pdb on atom "b"
    s.x <- s.chain$atom$b
  
#input 4: generate the plot
      plotb3(s.x, sse = s.chain, typ = "l", ylab = "Bfactor", main = (files))

}

```


## Final Output

```{r}

pdi_plot(files)

```

